---
title: "Visualizing Pi"
output:
  html_document: default
  html_notebook: default
---

We celebrate Pi-day (3/14) every year at the library. And though it is december, and there is months to the big day, it is not too early to begin planning. This is something we do in the incredibly amounts of time we have left over when all the other tasks that also only take a tiny portion of the day are done.
In other words, this is something we plan in our weekly lunchbreak.

Usually Henrik delivers a small talk about Pi. We eat some Pie, and there is much rejoicing.

But we would like to show something else. A cool visualization perhaps. There are several out there, but it is no fun just to show cool stuff. Its much cooler to understand how it is done.

One of the visualizaions we found on the net is this. https://www.visualcinnamon.com/portfolio/the-art-in-pi It is made by the extremely talented Nadieh Bremer. Take a look at her page. She does a lot of other cool stuff!

So. How to replicate this? 

The idea is to take pi. Draw a line segment from 0,0 in the plane to a new point, determined by the first digit. Then we draw another segment from that point, to a newer point, determined by the second digit in Pi. Continue ad nauseam, or until your computer catches fire. Add colours, and make it look cool.

First item is to get pi to a million decimal places.
Read it in, and make a vector containing all the digits.
I'll begin with a smaller string.
```{r}
numvec <- "3.1415926535897932384626433832795028841971693993751058209749445923078164062862089986280348253421170679"
```

Not quite pi. But I can't be bothered figuring out how many digits of pi I should use in order to get all the digits from 0 to 9. I want that to make sure that whatever I do, works on all the digits.

Next up is getting rid of the period. 
```{r}
numvec <- gsub("[[:punct:]]", "", numvec)

```

In Denmark we would use a "," instead of "." This removes all punctuation.

Now we convert it to a vector with individual digits. And cast them as integers rather than characters.
```{r}
numvec <- as.integer(unlist(strsplit(numvec,"")))

```

We begin a (0,0). The next point is determined by the first digit i pi - 3. The number 3 should determine witch direction, angle, we should move. The new x-value will be cos(f(3)), where f(3) is some function, that converts 3 to the correct angle. That should be some fraction of pi. The new y-value is similarly determined by sin(f(3)).

What should f(x) be? Imagine a decimal clock, with the ten digits 0 to 9 positioned around the circle. In the original 0 is at top, at 12 o'clock. 5 is at 6 o'clock. And the numbers follow the circle clockwise.
In the unity circle 0 is positioned at 1/2pi. 5 at 1 1/2 pi. So f(0)=1/2pi, f(5)=1 1/2pi. 

This formula gives us the result we need:
5/2*pi - z/5*pi.

z is the digit in pi. For all xof them, we can calculate the steps we need in this way:
```{r}
x <- cos(5/2*pi - numvec/5*pi)
y <- sin(5/2*pi - numvec/5*pi)
```

So. Beginning at (x0,y0), we calculate the trig functions, add it to (x0,y0) and get (x1,y1). When we continue to do that, we notice, that xi is the cumulative sum of all the calculated x-values. If we calculate the trig functions for all of the digits in pi, and then calculate the cumulative sum, we get all the x-values. Similarly with all the y-values.

Thats neat, and thanks to Henrik, who pointed this out. We get the cumulative sums easily:
And we can then get the cumulative sums with:
```{r}
x <- cumsum(x)
y <- cumsum(y)
```

Thats it! There is just a little detail. The first value in x,y is the first step we need to take - or the second point if you will.
We will need to add a 0 at the start:

```{r}
x <- c(0,x)
y <- c(0,y)
```

Thats it! I should now be able to plot it. Lets just do a quick test:

```{r}
plot(x,y, type="l")
```

That looks almost like what we need. The most notable difference is that Nadieh only plots the decimals. I include the "3" as well.
We are going to test this, and it is cumbersome to do all the preceding steps.
Let's begin with defining the generation of points as a function, that takes a string. In just a little while I'll invoke the magick of ggplot. So it would be convenient that the result is a dataframe:

```{r}
piPoints <- function(piString){
  numbers <- gsub("[[:punct:]]", "", piString)
  numbers <- as.integer(unlist(strsplit(numbers,"")))
  x <- cos(5/2*pi - numbers/5*pi)
  y <- sin(5/2*pi - numbers/5*pi)
  x <- cumsum(x)
  y <- cumsum(y)
  x <- c(0,x)
  y <- c(0,y)
  df <- data.frame(as.integer(c(0,numbers)),x,y)
  colnames(df) <- c("num", "x", "y")
  return(df)
}
```

All right. We now have a dataframe with three columns. The number, and the x,y coordinates. We can plot that with ggplot:
```{r}

df <- piPoints(numvec)
library(ggplot2)

ggplot(df, aes(x=x,y=y,group="1"))+
  geom_path(aes(colour=factor(df$num)))
```

Weird. Something is wrong. Look at the second linesegment. It goes up and to the right, just as it should, since this segment represents a "1". But the color? It is green. It should be sorta orangy. The reason is that ggplot looks at the number, and by implication the color, at the origin of the linesegment. Not at the end.
How to change that? What I really need to do, is shifting the number column one place relative to the x,y pairs.
Simply done - remove the 0 at the beginning of the num-vector in the function. And the final digits in the x and y vectors:
```{r}
piPoints <- function(piString){
  numbers <- gsub("[[:punct:]]", "", piString)
  numbers <- as.integer(unlist(strsplit(numbers,"")))
  x <- cos(5/2*pi - numbers/5*pi)
  y <- sin(5/2*pi - numbers/5*pi)
  x <- cumsum(x)
  y <- cumsum(y)
  x <- c(0,x)
  y <- c(0,y)
  df <- data.frame(as.integer(c(numbers)),x[-length(x)],y[-length(y)])
  colnames(df) <- c("num", "x", "y")
  return(df)
}
```
Lets take another look:
```{r}
df <- piPoints(numvec)
library(ggplot2)

ggplot(df, aes(x=x,y=y,group="1"))+
  geom_path(aes(colour=factor(df$num)))
```

Now we're there. Two steps are missing: Adding more digits - I would like to get to one million digits. And getting it to look nice. Lets begin with the last step.

I feel the need. The need for speed
==================
This is the writeup. So I did all this, without looking at the original code. In the original, the cumulative sums are calculated differently:

x <- y <- rep(NULL, length(numvec))
x[1] <- 0
y[1] <- 0

for (i in 2:length(piVec)){
    x[i] <- x[(i-1)] + sin((pi*2)*(numvec[i]/10))
    y[i] <- y[(i-1)] + cos((pi*2)*(numvec[i]/10))  
}

That is nice, and the way to do it that I was unable to figure out on my own. Because I got lost in apply-functions using lag(). Which is not the way to do it.

But is my way faster? I would guess it is, as for-loops in general are not very fast in R.

I'll use microbenchmark to do that. You will probably need to install it:
devtools::install_github("olafmersmann/microbenchmarkCore")
devtools::install_github("olafmersmann/microbenchmark")

Lets run the test:
```{r}
library(microbenchmark)
mbm <- microbenchmark("lm" ={
x <- cos(5/2*pi - numvec/5*pi)
y <- sin(5/2*pi - numvec/5*pi)
x <- cumsum(x)
y <- cumsum(y)
x <- c(0,x)
y <- c(0,y)
},{
x <- y <- rep(NULL, length(numvec))
x[1] <- 0
y[1] <- 0
for (i in 2:length(numvec)){
    x[i] <- x[(i-1)] + sin((pi*2)*(numvec[i]/10))
    y[i] <- y[(i-1)] + cos((pi*2)*(numvec[i]/10))  
}
}

)
mbm
```

Not a very nice output. Which doesn't change the fact that the way I do it (with an assist from Henrik), is quite a bit faster than the original.